Appendix

### 2024-09-11 manuscript/manuscript.Rmd
Updated real-world datasets analysis.

@@

A common feature in health-related datasets is age.
Here we use data on 228 patients with lung cancer that was collected and published by Loprinzi et al. [@Loprinzi1994-cd].
The age in the cohort was $62.4 \pm 9.1$ (mean ± standard deviation) years.
Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:lung-cancer-age}:
no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).

```{r lung-cancer-age, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Quantile-quantile plots for age of patients with lung cancer, expressed in years. ",
  "The left panel shows the quantile-quantile plot for the original feature ",
  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
  "The right panel shows the quantile-quantile plots for the feature transformed using ",
  "the non-robust and robust location- and-scale invariant transformations."
)

out <- .plot_lung_cancer_age(plot_theme = plot_theme)
out$plot
```

@@

Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:penguin-body-mass}:
no transformation (residual sum $\sum r_i = 48.0$);
conventional transformation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.102$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).

```{r penguin-body-mass, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Quantile-quantile plots for penguin body mass, measured in grams. ",
  "The left panel shows the quantile-quantile plot for the original feature. ",
  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
  "The right panel shows the quantile-quantile plots for the feature transformed using ",
  "the non-robust and robust location- and-scale invariant transformations."
)

out <- .plot_penguin_body_mass(plot_theme = plot_theme)
out$plot
```
Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).

@@

Geospatial datasets usually contain coordinates.
The Ames housing dataset contains data on 2930 properties that were sold between 2006 and 2010 [@De-Cock2011-jf], including their geospatial coordinates.
The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:ames-latitude}:
no transformation (residual sum $\sum r_i = 328$);
conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$);
location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$).

```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Quantile-quantile plots for latitude coordinate in a housing dataset. ",
  "The left panel shows the quantile-quantile plot for the original feature ",
  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
  "The right panel shows the quantile-quantile plots for the feature transformed using ",
  "the non-robust and robust location- and-scale invariant transformations."
)

out <- .plot_ames_latitude(plot_theme = plot_theme)
out$plot
```
Every transformation reduced the residual sum.
The non-robust location- and scale-invariant transformation did not improve over conventional alternatives and yielded a data distribution lacking central normality (empirical central normality test: $p=0.05$).
However, conventional transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.

@@

@@

```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Quantile-quantile plots for fuel consumption data in the Top Gear dataset. ",
  "The feature contains three strong outliers related to highly fuel-efficient vehicles.",
  "The left panel shows the quantile-quantile plot for the original feature. ",
  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.76$).",
  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.55$, respectively. ",
  "The right panel shows the quantile-quantile plots for the feature transformed using ",
  "the non-robust and robust location- and scale-invariant transformations. ",
  "For these, the empirical central normality test yields $p=0.03$, and $p=0.86$, respectively. ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

out <- .plot_top_gear(
  plot_theme = plot_theme
)

out$plot
```

The Top Gear dataset contains data on 297 vehicles that appeared on the BBC television show *Top Gear*. Within this dataset, the fuel consumption feature contains outliers due to highly fuel-efficient vehicles. Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively). Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

@@

```{r ischemic-stroke, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Quantile-quantile plots for the maximum arterial wall thickness in patients with ischemic stroke. ",
  "This feature contains two strong outliers. ",
  "The left panel shows the quantile-quantile plot for the original feature. ",
  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.56$). ",
  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
  "For these, the empirical central normality test yields $p=0.01$, and $p=0.61$, respectively. ",
  "The right panel show the quantile-quantile plots for the feature transformed using ",
  "the non-robust and robust location- and scale-invariant transformations. ",
  "For these, the empirical central normality test yields $p=0.02$, and $p=0.10$, respectively. ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

out <- .plot_ischemic_stroke(
  plot_theme = plot_theme
)

out$plot
```
The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
Angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
The maximum arterial wall thickness feature contains several instances with outlier values.
Figure \ref{fig:ischemic-stroke} show that non-robust transformation fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively). Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

@@

We used 285 datasets from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
In this collection, 122 datasets correspond to regression tasks and 163 datasets
to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
each dataset was used to train a model for each of 16 process configurations.
Each process configuration specifies the learner (generalised linear model or
random forest), transformation method (none, conventional Yeo-Johnson, robust
invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
normality test ($p = 0.01$)), and normalisation method (none,
$z$-standardisation), yielding 16 distinct configurations.
Before each experiment, each dataset was randomly split into a training (70%)
and holdout test (30%) set five times. Thus, a total of 22800 models were
created. Each model was then evaluated using the holdout test set using one of
two metrics, i.e. the root relative squared error (RRSE) for regression tasks and the
area under the receiver operating characteristic curve (AUC) for classification tasks.

@@

This study had the following limitations.
First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-C).

>>>>

```{r experimental-results-invariance, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=5.0}
cap <- paste0(
  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
  "penguin body mass (middle row); ",
  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Yeo-Johnson transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Yeo-Johnson transformations (right column)."
)

data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "yeo_johnson")
data$p
```

@@

A common feature in health-related datasets is age.
Here we use data on 228 patients with lung cancer that was collected and published by Loprinzi et al. [@Loprinzi1994-cd].
The age in the cohort was $62.4 \pm 9.1$ (mean ± standard deviation) years.
Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).

@@

Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
no transformation (residual sum $\sum r_i = 48.0$);
conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).

Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).

@@

Geospatial datasets usually contain coordinates.
The Ames housing dataset contains data on 2930 properties that were sold between 2006 and 2010 [@De-Cock2011-jf], including their geospatial coordinates.
The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
no transformation (residual sum $\sum r_i = 328$);
conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$);
location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$).

Every transformation reduced the residual sum.
The non-robust location- and scale-invariant transformation did not improve over conventional alternatives and yielded a data distribution lacking central normality (empirical central normality test: $p=0.05$).
However, conventional transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.

@@

```{r experimental-results-outlier-robustness, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.5}
cap <- paste0(
  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Yeo-Johnson transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Yeo-Johnson transformations (right column). ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "yeo_johnson")
data$p
```

@@

The Top Gear dataset contains data on 297 vehicles that appeared on the BBC television show *Top Gear* [@Alfons2021-kc].
Within this dataset, the fuel consumption feature contains outliers due to highly fuel-efficient vehicles.

Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
no transformation (residual sum $\sum r_i = 110$, $p=0.76$);
conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.01$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 137$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.61$);
location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.13$);
and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.10$).

Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively).
Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

@@

The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
Angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
The maximum arterial wall thickness feature contains several instances with outlier values.
Figure \ref{fig:experimental-results-outlier-robustness} shows that non-robust transformations fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

@@

We used 285 datasets from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
In this collection, 122 datasets correspond to regression tasks and 163 datasets
to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
each dataset was used to train a model for each of 16 process configurations.
Each process configuration specifies the learner (generalised linear model or
random forest), transformation method (none, conventional Yeo-Johnson, robust
invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
normality test (rejecting transformations with $p \leq 0.01$), and normalisation method (none,
$z$-standardisation), yielding 16 distinct configurations.
Before each experiment, each dataset was randomly split into a training (70%)
and holdout test (30%) set five times. Thus, a total of 22800 models were
created. Each model was then evaluated using the holdout test set using one of
two metrics, i.e. the root relative squared error (RRSE) for regression tasks and the
area under the receiver operating characteristic curve (AUC) for classification tasks.

@@

As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (appendix B).

### 2024-09-11 manuscript/manuscript_appendix.Rmd
Added plotting function for outlier data.
Updated real-world datasets analysis.

```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Blabla"
)

@@

>>>>

The effect of using location- and scale-invariant transformations was investigated using real-world datasets.

## Invariance

Results for Box-Cox transformations of features without outliers are shown in Figure \ref{fig:experimental-results-invariance-appendix}.

```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
  "penguin body mass (middle row); ",
  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Box-Cox transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Box-Cox transformations (right column)."
)

@@

### Age of patients with lung cancer

Applying conventional and invariant Box-Cox transformations to age of patients with lung cancer [@Loprinzi1994-cd] yielded the following results:
no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$);
and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$).

Compared to location- and scale-invariant Yeo-Johnson transformations, the Box-Cox transformations do not reduce residuals compared to conventional variants.

### Penguin body mass

Applying conventional and invariant Box-Cox transformations to the body mass of penguins [@Gorman2014-eo] yielded the following results:
no transformation (residual sum $\sum r_i = 48.0$);
conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$);
and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$).

Just as for location- and scale-invariant Yeo-Johnson transformations, Box-Cox transformations produced a lower overall residual sum compared to their conventional counterparts.
Similarly, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.

### Latitude in the Ames housing dataset

Applying conventional and invariant Box-Cox transformations to the latitude of houses in the Ames housing dataset [@De-Cock2011-jf] yielded the following results:
no transformation (residual sum $\sum r_i = 328$);
conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$);
location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$);
and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$.

Similar to conventional Yeo-Johnson transformations (non-robust and robust), conventional Box-Cox transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
Location- and scale-invariant Box-Cox transformations did not suffer from this issue.

## Robustness against outliers

Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:results-outlier-robustness-appendix}.

```{r experimental-results-outlier-robustness-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Box-Cox transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Box-Cox transformations (right column). ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "box_cox")
data$p
```

### Fuel efficiency in the Top Gear dataset

The Top Gear dataset contains data on 297 vehicles, with outliers related highly fuel-efficient vehicles [@Alfons2021-kc].
Applying conventional and invariant Box-Cox transformations to the fuel consumption feature yielded the following results:
no transformation (residual sum $\sum r_i = 54$, $p=0.76$);
conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p=0.01$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.55$);
location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{BC} = 0.5$, $\sigma_{BC} = 0.1$, $p=0.03$);
and robust location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 49$, $\mu_{BC} = 2.0$, $\sigma_{BC} = 2.3$, $p=0.58$).

### Maximum arterial wall thickness in an ischemic stroke dataset

The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
Applying conventional and invariant Box-Cox transformations to the maximum arterial wall thickness feature yielded the following results:
no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.1$, $p=0.01$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{BC} = 7.2$, $\sigma_{BC} = 14$, $p=0.61$);
location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{BC} = -11.8$, $\sigma_{BC} = 6.9$, $p=0.13$);
and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.10$).
